\begin{itemize}

\item{{ALTAZ} --- Move the telescope to a specified alt/az position}

\item{{CD} --- {\bf C}lose the {\bf D}ome slit.}
\item{{CENTER} --- move the telescope to {\bf CENTER} on an object in an image based on the current image display.}
\item{{CF} --- {\bf C}lose the command {\bf F}ile.}
\item{{CL} --- {\bf C}lose the {\bf L}ouvers - the large vents at the floor of the dome.}
\item{{CLOSE} --- {\bf CLOSE}s the guide shutter.}
%\item{{CM} --- {\bf C}lose the {\bf M}irror covers.}
\item{{CO} --- move the telescope to the {\bf CO}ordinates given as hh mm ss dd mm ss.}

\item{{DI} --- {\bf D}ome {\bf I}nitialization command.}
\item{{-DISK/+DISK} --- turn off/on autosaving of images to {\bf DISK}.}
\item{{-DISPLAY/+DISPLAY} --- turn off/on auto {\bf DISPLAY}ing of images.}
\item{{DF [{\it $\Delta$f}]} --- move focus position by {\bf D}elta {\bf F}ocus = {\it $\Delta$f}.}
\item{{DM} --- toggle {\bf D}o{\bf M}e tracking on/off.}

\item{{EXP [{\it time}]} --- take an {\bf EXP}osure of length {\it time} seconds.}

\item{{FILENAME [{\it name}]} --- change the default root for image {\bf FILENAME}s to {\it name}.}
\item{{FILL} --- {\bf FILL} the dewar}
%\item{{FILTER [{\it n}]} --- switch current {\bf FILTER} to position {\it n}.}
%\item{{FITS} --- set file type to {\bf FITS} format and turn on autosaving of images.}
\item{{FLAT {\it ndesired nmax startexp fudge}} --- takes a series of {\bf FLAT} field images; see text for more.}
\item{{FO [{\it f}]} --- move to {\bf FO}cus position {\it f}.}
\item{{FOCRUN} --- {\bf FO}cus {\bf RUN}.}
\item {{FUDGE} --- setup for STAN command.}
\item{{FULLSCALE} --- set autoscaling of images to have greyscaling showing {\bf FULLSCALE} of image.}

\item{{GCENTER} --- CENTER command for guider.}
\item{{GEXP} --- EXP command for guider.}
\item{{GFLAT} --- FLAT command for guider.}
\item{{GFOCRUN} --- {\bf FO}cus {\bf RUN} in the {\bf G}uider.}
\item{{GFULLSCALE} --- FULLSCALE command for guider.}
\item{{GLOCATE} --- LOCATE command for guider.}
\item{{GOFFSET [$\Delta${\it x} $\Delta${\it y}]} --- {\bf Offset} the telescope  by a specified number of \textit{pixels} in the guider}
\item{{GSAMESCALE} --- SAMESCALE command for guider.}
%\item{{GSSCALE} --- SSCALE command for guider.}
\item{{GSKYSCALE} --- SKYSCALE command for guider.}
\item{{GUIDE} --- turn on the {\bf GUIDE} system.}
\item{{GUIDEOFF} --- turn the {\bf GUIDE} system {\bf OFF}; for long exposures, a warning will come up if guiding if off.}
\item{{GUPDATE [{\it n}]} --- {\bf UPDATE} the number of {\bf G}uide exposures to determine centroid.}
\item{{GSIZE [{\it n}]} --- set the {\bf SIZE} of box used by {\bf G}uide camera for computing star centroids.}

%\item{{HO} --- initialize telescope to {\bf HO}me position.}
\item{{HP} --- {\bf H}el{\bf P} command.}

\item{{INIT} --- INITialize telescope.}
\item{{INPUT} filenmae --- execute a series of commands from the specified
{\bf INPUT} file}
\item{{INST} --- select a new default instrument for pointing (currently, either
science CCD or guider CCD.}

\item{{LOCATE} --- move the telescope such that object in current image display
 is moved to some other specified {\bf LOCATION} in the display.}

%\item {{MAG} --- setup for STAN command.}
\item{{MEXP [{\it time number}]} --- take {\it number} of {\bf M}ultiple 
{\bf EXP}osures of length {\it time} seconds.}

\item{{NE} --- change to a {\bf N}ew {\bf E}poch for coordinates.
\item{{NEWEXT [\it ext}]} --- give a {\bf NEW} {\bf EXT}ension for the next image file.}
\item{{NOSHUTTER} --- set guide camera to use {\bf NO} {\bf SHUTTER} during guide exposures.} 

\item{{OBJECT} --- specify {\bf OBJECT} name for FITS header.}
\item{{OD} --- {\bf O}pen the {\bf D}ome slit.}
\item{{OF} --- {\bf O}pen the command {\bf F}ile created by user.}
\item{{OFFSET [$\Delta${\it x} $\Delta${\it y}]} --- {\bf Offset} the telescope  by a specified number of \textit{pixels}.}
%\item{{OM} --- {\bf O}pen the {\bf M}irror covers.}
\item{{OL} --- {\bf O}pen the {\bf L}ouvers - the large vents at the floor of the dome.}
\item{{OPEN} --- {\bf OPEN} the guide camera shutters.}
\item{{OS} --- {\bf O}pen the {\bf S}tandard star file.}

\item{{PA} --- change the {\bf P}osition {\bf A}ngle of the detector.}
\item{{PM} --- move the telescope to a position of one of the last five
{\bf P}revious {\bf M}ove positions.}

\item{{QCK [{\it time}]} --- take a {\bf Q}ui{\bf CK} exposure of 
length {\it time} seconds, \textbf{but do not save it}.}
\item{{QM [{\it $\Delta$ RA $\Delta$ DEC }]} --- do a {\bf Q}uick {\bf M}ove with the offset in RA and DEC in arcsec.}
\item{{QU} --- {\bf QU}it the telescope command program {\it gracefully}.}

\item{{RF [{\it id}]} --- {\bf R}ead from command {\bf F}ile created by user and move telescope to object {\it id}.}
\item{{RS} --- {\bf R}ead from {\bf S}tandard star file.}

\item{{SA} --- set the telescope to point at a {\bf SA}o star.}
\item{{SAMESCALE} --- set autoscaling of images for future images to be the {\bf SAMESCALE} as current image.}
\item{{SCALE {\it low high}} --- redisplay current image with new grey {\bf SCALE} between {\it low} and {\it high}.}
\item{{SHUTTER} --- sets guide CCD camera into normal {\bf SHUTTER} mode.}
\item{{SKYSCALE} --- set autoscaling of images to have grey{\bf SCALE} levels around mean {\bf SKY} level.}
\item{{SLEEP [{\it t}]} --- suspend operation (sleep) for specified number of seconds.}
\item{{STAN [{\it n}]} --- take an exposure through filter number {\it n} of a {\bf STAN}dard star; program determines exp time.}
\item{{ST} --- {\bf ST}ow the telescope to stow position to quit.}

\item{{TR} --- toggle {\bf TR}acking on and off.}

\item{{UC} -- {\bf U}pdate {bf C}oordinates for current telescope pointing.}

\item{{WRITE} --- toggles to mode where guide exposures are {\bf WRITE}en to disk; default is not to save images.}

%\item {{ZERO} --- setup for STAN command.}
\end{itemize}
